Parsing Postfix log files

### 2007-07-11 doc/logparser.tex
Added support for generic rules which will be tried after program specific rules. Documented generic rules. Changed all the LDAP rules to be generic rules.

    \item [program] The program (postfix/smtp, postfix/smtpd, postfix/qmgr,
        etc.) whose log lines the rule applies to.  This avoids needlessly
        trying rules which won't match the line, or worse, might match
        unintentionally.

@@

    \item 7 mails (23 warnings) have out of order log entries, leading to
        warnings about missing data and queueids.  The two log files where
        the out of order log lines occur are more than an order of
        magnitude larger than normal, due to the extremely large volume of
        mail handled by the server on those days; process scheduling
        vagaries and the undefined order of messages sent to syslog(8) may
        explain why the lines were out of order.

@@

The first of these problems cannot be dealt with; although out of order log
lines (the second problem) are dealt with in other circumstances (see
sections~\ref{tracking-re-injected-mail},~\ref{discarding cleanup
lines}~and~\ref{pickup logging after cleanup}), dealing with these lines
would introduce a large amount of complication (deferred warnings, deferred
committal, ambiguity) for a problem which occurs only seven times in
60721709 log lines.

The state tables contain entries for mails not yet delivered when the
parser finishes execution.  Ideally all they should contain are mails which
are awaiting delivery after the period covered by the logs, though they may
also contain mails whose initial entries are not contained in the logs.
Any other entries are evidence of a failure in parsing or an aberration in
the logs.  After parsing the 93 test log files the state tables contain 27
entries, breaking down into:

@@

    \item 2 connections which started only seconds before the logs ended
        and had not completed.

@@

    \item 5 mails resulting from the out of order log entries described
        above.

@@

    \item It appears that the hostname used in the HELO command is not
        logged if the mail is accepted.\footnote{Tested with Postfix 2.2.10
        and 2.3.7; this may possibly have changed in Postfix 2.4.}  It
        should be reasonably simple to write a policy server which causes
        Postfix to log a warning containing the HELO hostname when the DATA
        command is accepted.

@@

        XXX EXPAND ON THIS

>>>>

    \item [program] The program (postfix/smtp, postfix/smtpd, postfix/qmgr,
        etc.) whose log lines the rule applies to.  This avoids needlessly
        trying rules which won't match the line, or worse, might match
        unintentionally.  Rules whose program is \texttt{*} will be tried
        against any lines which aren't parsed by program specific rules.

@@

    \item 5 mails (20 warnings) have out of order log entries, leading to
        warnings about missing data and queueids.  The two log files where
        the out of order log lines occur are more than an order of
        magnitude larger than normal, due to the extremely large volume of
        mail handled by the server on those days; process scheduling
        vagaries and the undefined order of messages sent to syslog(8) may
        explain why the lines were out of order.

@@

The first of these problems cannot be dealt with; although out of order log
lines (the second problem) are dealt with in other circumstances (see
sections~\ref{tracking-re-injected-mail},~\ref{discarding cleanup
lines}~and~\ref{pickup logging after cleanup}), dealing with these lines
would introduce a large amount of complication (deferred warnings, deferred
committal, ambiguity) for a problem which occurs only five times in
60721709 log lines.

The state tables contain entries for mails not yet delivered when the
parser finishes execution.  Ideally all they should contain are mails which
are awaiting delivery after the period covered by the logs, though they may
also contain mails whose initial entries are not contained in the logs.
Any other entries are evidence of a failure in parsing or an aberration in
the logs.  After parsing the 93 test log files the state tables contain 25
entries, breaking down into:

@@

    \item 1 connection which started only seconds before the logs ended
        and had not completed.

@@

    \item 4 mails resulting from the out of order log entries described
        above.

@@

    \item It appears that the hostname used in the HELO command is not
        logged if the mail is accepted.\footnote{Tested with Postfix 2.2.10
        and 2.3.11; this may possibly have changed in Postfix 2.4.}  It
        should be reasonably simple to write a policy server which causes
        Postfix to log a warning containing the HELO hostname when the DATA
        command is accepted.

@@
